---
title: "explore-historical-data (experimental)"
author: "Daniel Moul"
date: "`r Sys.Date()`"
output: html_document
---

<br><br>

***This is an experimental script used prior to the development of nc-camping-writeup.Rmd. It is not needed and may not work correctly.***

<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse, warn.conflicts = FALSE)
library(arrow, warn.conflicts = FALSE)
library(jsonlite)
library(janitor)
library(gt)
library(scales)
library(glue)
library(lubridate)
library(purrr)

library(hrbrthemes)
# remotes::install_github("hrbrmstr/hrbrthemes")
# run 2021-08-01 to address segfault errors like this:
# sh: line 1: 11020 Segmentation fault: 11  '/Library/Frameworks/R.framework/Versions/4.1/Resources/library/Rttf2pt1/exec//ttf2pt1' -a -GfAe '/Library/Frameworks/R.framework/Versions/4.1/Resources/library/hrbrthemes/fonts/roboto-condensed/RobotoCondensed-Bold.ttf' '/var/folders/0j/hjw8sflx19l6rw4s1l68qq640000gn/T//RtmpGN3PSJ/fonts/RobotoCondensed-Bold' 2>&1
theme_set(theme_ipsum_ps() + 
            theme(panel.grid.minor = element_blank()))
import_plex_sans()

# library(readr)
# options(readr.read_lazy = TRUE)

options(dplyr.summarise.inform = FALSE)

big_campground_size <- 50 # fifty or more campsites

```

```{r}
facilities_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/Facilities_API_v1.json",
                      flatten = FALSE)

facilities <- facilities_raw[[1]] %>%
  mutate_if(is.character, ~map_chr(.x, iconv, "UTF-8", "UTF-8", sub='')) %>% # just in case
  clean_names() %>%
  mutate(facility_name = str_to_title(facility_name),
         facility_name = if_else(facility_id == "250838", "Halfway Group Campground", facility_name)
         ) %>%
  select(facility_id, facility_name, facility_type_description)


```


```{r}
nc_hist <- read_feather("./data/processed/nc_historical_camping.feather") %>%
  mutate_if(is.character, ~map_chr(.x, iconv, "UTF-8", "UTF-8", sub='')) %>% # just in case
  left_join(.,
            facilities ,
            by = "facility_id") %>%
  filter(!is.na(path),
         use_type == "Overnight") %>%
  mutate(year = year(start_date),
         park = str_to_title(park),
         park = case_when(
           str_detect(park, "Doughton Park")        ~ "Doughton Park Campground",
           str_detect(park, "Julian Price")         ~ "Julian Price Campground",
           str_detect(park, "Linville Falls")       ~ "Linville Falls Campground",
           str_detect(park, "Cove Creek")           ~ "Cove Creek Campgrounds",
           str_detect(park, "Badin Lake")           ~ "Badin Lake Campgrounda",
           str_detect(park, "Cataloochee")          ~ "Cataloochee Campgrounds",
           str_detect(park, "Deep Creek")           ~ "Deep Creek Campground",
           str_detect(park, "Big Creek")            ~ "Big Creek Campgrounds",
           str_detect(park, "Smokemont")            ~ "Smokemont Campgrounds",
           str_detect(park, "White Pine")           ~ "White Pine Campgrounds",
           str_detect(park, "Cheoah Point")         ~ "Cheoah Point Campgrounds",
           TRUE                                     ~ park
         ),
         path_year = as.numeric(str_extract(path, "\\d+")),
         site_type_simple = case_when(
           str_detect(site_type, "TENT")        ~ "tent",
           str_detect(site_type, "CABIN")       ~ "cabin",
           str_detect(site_type, "GROUP")       ~ "group",
           str_detect(site_type, "RV")          ~ "rv",
           str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
           str_detect(site_type, "STANDARD")    ~ "standard",
           str_detect(site_type, "SHELTER")     ~ "shelter",
           str_detect(site_type, "MANAGEMENT")  ~ "management",
           TRUE                                 ~ site_type
         )
  )

```

How many reservations per year?

```{r}
nc_hist %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  adorn_totals(where = "row") %>%
  gt() %>%
  fmt_number(columns = c(n),
             decimals = 0)
```

<br><br>

Well, 2006 and 2018 are missing. But otherwise looks ok.

<br>

What if I group by file instead of a date field?

<br>

```{r}
nc_hist %>%
  group_by(path) %>%
  summarize(n = n()) %>%
  adorn_totals(where = "row") %>%
  gt() %>%
  fmt_number(columns = c(n),
             decimals = 0)
```

<br><br>

How about total camping nights per year?

<br>

```{r}
nc_camping_nights <- nc_hist %>%
  #mutate(year = as.numeric(str_extract(path, "\\d+"))) %>%
  group_by(year) %>%
  summarize(n = n(), 
            n_nights = sum(nights, na.rm = TRUE)) 

nc_camping_nights %>%
  adorn_totals(where = "row")%>%
  gt() %>%
  fmt_number(columns = c(n, n_nights),
             decimals = 0)

```

<br><br>



```{r}
normalization <- nc_camping_nights %>%
  filter(year == 2008) %>%
  pull(n_nights)

nc_camping_nights_for_plot <- nc_camping_nights %>%
  filter(year >= 2007) %>% 
  mutate(nights_normalized = n_nights / normalization) 

nc_camping_nights_for_plot %>%
  ggplot(aes(year, nights_normalized)) +
  geom_line() +
  geom_point(size = 3) +
  #scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = percent_format()) +
  expand_limits(y = 0, x = 2005) +
  labs(title = "Person-nights of camping in NC",
       subtitle = glue("Pecentage of 2008 person-nights; 2008 = {comma(normalization)}"),
       x = "",
       y = "")

```

<br>

There is a lot of missing data, especially in the earlier years. Let's look at the same graph but include only campgrounds that had data in 2008:

```{r}

nc_camping_nights <- nc_hist %>%
  mutate(year = as.numeric(str_extract(path, "\\d+"))) %>%
  group_by(year, facility_name) %>%
  summarize(n = n(), 
            n_nights = sum(nights, na.rm = TRUE)) %>%
  ungroup()

campgrounds_with_2008_data <- nc_camping_nights %>%
  filter(year == 2008) %>%
  select(facility_name)

normalization <- nc_camping_nights %>%
  filter(year == 2008) %>%
  pull(n_nights) %>%
  sum()

nc_camping_nights_for_plot <- nc_camping_nights %>%
  inner_join(., campgrounds_with_2008_data,
             by = "facility_name") %>%
  group_by(year) %>%
  summarize(n = n(), 
            n_nights = sum(n_nights, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(year >= 2007) %>% 
  mutate(nights_normalized = n_nights / normalization) 

nc_camping_nights_for_plot %>%
  ggplot(aes(year, nights_normalized)) +
  geom_line() +
  geom_point(size = 3) +
  #scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = percent_format()) +
  expand_limits(y = 0, x = 2005) +
  theme(plot.title.position = "plot") +
  labs(title = "Accelerating growth in camping person-nights in NC",
       subtitle = glue("Pecentage of 2008 person-nights; 2008 = {comma(normalization)}",
                       "\nn = {nrow(campgrounds_with_2008_data)}; excludes campsites lacking 2008 data"),
       x = "",
       y = "")
```


<br>

Have there been changes in the type of campsites?

```{r}
nc_camping_nights_site_type <- nc_hist %>%
  group_by(year, site_type_simple) %>%
  summarize(n = n(), 
            n_person_nights = sum(people_nights, na.rm = TRUE)) %>%
  ungroup()

nc_camping_nights_site_type %>%
  gt() %>%
  fmt_number(columns = n_person_nights,
             decimals= 0)
```


```{r eval=FALSE}
normalization <- nc_camping_nights %>%
  filter(year == 2008) %>%
  pull(n_nights)
```


```{r}
nc_camping_nights_for_plot <- nc_camping_nights_site_type %>%
  filter(year >= 2008,
         year != 2018) #%>% 
  # mutate(nights_normalized = n_nights / normalization) 

nc_camping_nights_for_plot %>%
  # since this isn't working, I'll do it manually: fct_lump(site_type_simple, n = 5, other_level = "other")
  mutate(site_type_simple =  case_when(
    site_type_simple == "shelter"     ~ "equest+other",
    site_type_simple == "management"  ~ "equest+other",
    site_type_simple == "equestrian"  ~ "equest+other",
    TRUE                              ~ site_type_simple
    ),
    site_type_simple = fct_reorder(site_type_simple, n_person_nights, sum)
  ) %>%
  group_by(year, site_type_simple) %>%
  summarize(n_person_nights = sum(n_person_nights)) %>%
  ungroup() %>%
  ggplot(aes(year, n_person_nights, color = site_type_simple)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = label_number_si()) +
  scale_color_viridis_d(end = 0.92) +
  guides(color = guide_legend(reverse = TRUE)) +
  # expand_limits(y = 0, x = 2005) +
  labs(title = "Person-nights of camping in NC",
       #subtitle = glue("Pecentage of 2008 person-nights; 2008 = {comma(normalization)}"),
       x = "",
       y = "")

```

<br>

Whatever a "standard" camp site is ... that's where the action is.

The non-electric sites in particular:

```{r}
nc_camping_nights_site_type <- nc_hist %>%
  mutate(year = as.numeric(str_extract(path, "\\d+"))) %>%
  group_by(year, site_type) %>%
  summarize(n = n(), 
            n_person_nights = sum(people_nights, na.rm = TRUE)) %>%
  ungroup()

nc_camping_nights_for_plot <- nc_camping_nights_site_type %>%
  filter(year >= 2008) #%>% 
  # mutate(nights_normalized = n_nights / normalization) 

nc_camping_nights_for_plot %>%
  # since this isn't working, I'll do it manually: fct_lump(site_type_simple, n = 5, other_level = "other")
  mutate(
    # site_type_simple =  case_when(
    # site_type_simple == "shelter"     ~ "equest+other",
    # site_type_simple == "management"  ~ "equest+other",
    # site_type_simple == "equestrian"  ~ "equest+other",
    # TRUE                              ~ site_type_simple
    # ),
    site_type = fct_reorder(site_type, n_person_nights, sum)
  ) %>%
  filter(str_detect(site_type, "STANDARD|TENT|GROUP(.*)NONELECTRIC|RV"),
         #!str_detect(site_type, "GROUP")
         ) %>%
  group_by(year, site_type) %>%
  summarize(n_person_nights = sum(n_person_nights)) %>%
  ungroup() %>%
  ggplot(aes(year, n_person_nights, color = site_type)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = label_number_si()) +
  scale_color_viridis_d(end = 0.92) +
  guides(color = guide_legend(reverse = TRUE)) +
  # expand_limits(y = 0, x = 2005) +
  labs(title = "Person-nights of camping in NC",
       #subtitle = glue("Pecentage of 2008 person-nights; 2008 = {comma(normalization)}"),
       x = "",
       y = "")

```

<br>

Which campsites offer "standard" sites?


```{r}
nc_camping_nights_site_type <- nc_hist %>%
  filter(str_detect(site_type, "STANDARD") & !str_detect(site_type, "GROUP")) %>%
  mutate(year = as.numeric(str_extract(path, "\\d+"))) %>%
  group_by(year, site_type) %>%
  summarize(n = n(), 
            n_person_nights = sum(people_nights, na.rm = TRUE)) %>%
  ungroup()

nc_camping_nights_for_plot <- nc_camping_nights_site_type %>%
  filter(year >= 2008) #%>% 
  # mutate(nights_normalized = n_nights / normalization) 

nc_camping_nights_for_plot %>%
  # since this isn't working, I'll do it manually: fct_lump(site_type_simple, n = 5, other_level = "other")
  mutate(
    # site_type_simple =  case_when(
    # site_type_simple == "shelter"     ~ "equest+other",
    # site_type_simple == "management"  ~ "equest+other",
    # site_type_simple == "equestrian"  ~ "equest+other",
    # TRUE                              ~ site_type_simple
    # ),
    site_type = fct_reorder(site_type, n_person_nights, sum)
  ) %>%
  filter(str_detect(site_type, "STANDARD|TENT|GROUP(.*)NONELECTRIC|RV"),
         #!str_detect(site_type, "GROUP")
         ) %>%
  group_by(year, site_type) %>%
  summarize(n_person_nights = sum(n_person_nights)) %>%
  ungroup() %>%
  ggplot(aes(year, n_person_nights, color = site_type)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = label_number_si()) +
  scale_color_viridis_d(end = 0.92) +
  guides(color = guide_legend(reverse = TRUE)) +
  expand_limits(y = 0) +
  labs(title = "Person-nights of camping in NC",
       #subtitle = glue("Pecentage of 2008 person-nights; 2008 = {comma(normalization)}"),
       x = "",
       y = "")
```

<br>

```{r}
nc_camping_nights_site_type <- nc_hist %>%
  filter(str_detect(site_type, "STANDARD") & !str_detect(site_type, "GROUP")) %>%
  mutate(year = as.numeric(str_extract(path, "\\d+"))) %>%
  group_by(year, facility_name, site_type) %>%
  summarize(n = n(), 
            n_person_nights = sum(people_nights, na.rm = TRUE)) %>%
  ungroup()

nc_camping_nights_for_plot <- nc_camping_nights_site_type %>%
  filter(year >= 2008) #%>% 
  # mutate(nights_normalized = n_nights / normalization) 

nc_camping_nights_campground_for_plot <- nc_camping_nights_for_plot %>%
  # since this isn't working, I'll do it manually: fct_lump(site_type_simple, n = 5, other_level = "other")
  mutate(
    # site_type_simple =  case_when(
    # site_type_simple == "shelter"     ~ "equest+other",
    # site_type_simple == "management"  ~ "equest+other",
    # site_type_simple == "equestrian"  ~ "equest+other",
    # TRUE                              ~ site_type_simple
    # ),
    site_type = fct_reorder(site_type, n_person_nights, sum)
  ) %>%
  filter(str_detect(site_type, "STANDARD|TENT|GROUP(.*)NONELECTRIC|RV"),
         #!str_detect(site_type, "GROUP")
         ) %>%
  group_by(year, facility_name, site_type) %>%
  summarize(n_person_nights = sum(n_person_nights)) %>%
  ungroup() %>%
  group_by(year, facility_name,) %>%
  mutate(combined_n_person_nights = sum(n_person_nights)) %>%
  ungroup() %>%
  arrange(year, facility_name) %>%
  group_by(facility_name) %>%
  mutate(
    max_n = max(combined_n_person_nights),
    min_n = min(combined_n_person_nights),
    pct_change = last(combined_n_person_nights) / first(combined_n_person_nights)
         ) %>%
  ungroup()
  
```


```{r fig.height=8}
facility_name_levels <- nc_camping_nights_campground_for_plot %>%
  group_by(facility_name) %>%
  filter(year == max(year)) %>%
  ungroup() %>%
  distinct(facility_name, combined_n_person_nights) %>%
  arrange(desc(combined_n_person_nights)) %>%
  pull(facility_name)

nc_camping_nights_campground_for_plot %>%
  mutate(facility_name = factor(facility_name, levels = facility_name_levels),
         facility_name_label = glue("{facility_name}: {percent(pct_change, accuracy = 1)}"),
         facility_name_label = fct_reorder(facility_name_label, -combined_n_person_nights) #pct_change)
         ) %>%
  ggplot(aes(year, n_person_nights, color = site_type)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = label_number_si()) +
  scale_color_viridis_d(end = 0.92) +
  guides(color = guide_legend(reverse = TRUE)) +
  expand_limits(y = 0) +
  facet_wrap(~ facility_name, scales = "free_y") +
  labs(title = "Person-nights of camping in NC",
       subtitle = "Subplots ordered by decending number of person-nights",
       x = "",
       y = "")

```


<br>
<br>


```{r eval=FALSE}
nc_2020 <- read_feather("./data/processed/nc_by_year_semiprocessed/nc_2020.feather")

count(nc_2020, site_type)
```

Costs per person per night over time

```{r}
nc_camping_cost <- nc_hist %>%
  #filter(str_detect(site_type, "STANDARD") & !str_detect(site_type, "GROUP")) %>%
  filter(!str_detect(site_type, "MANAGEMENT")) %>%
  # mutate(year = as.numeric(str_extract(path, "\\d+")),
  #        site_type_simple = case_when(
  #          str_detect(site_type, "TENT")        ~ "tent",
  #          str_detect(site_type, "CABIN")       ~ "cabin",
  #          str_detect(site_type, "GROUP")       ~ "group",
  #          str_detect(site_type, "RV")          ~ "rv",
  #          str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
  #          str_detect(site_type, "STANDARD")    ~ "standard",
  #          str_detect(site_type, "SHELTER")     ~ "shelter",
  #          str_detect(site_type, "MANAGEMENT")  ~ "management",
  #          TRUE                                 ~ site_type
  #          )
  #        ) %>%
  group_by(year, site_type_simple) %>%
  summarize(n = n(), 
            n_person_nights = sum(people_nights, na.rm = TRUE),
            cost_median = median(paid_person_night, na.rm = TRUE),
            cost_avg = mean(paid_person_night, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
nc_camping_cost %>%
  filter(site_type_simple != "shelter") %>%
  mutate(site_type_simple = fct_reorder(site_type_simple, cost_median, last)) %>%
 ggplot(aes(year, cost_median, color = site_type_simple)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_viridis_d(end = 0.92) +
  guides(color = guide_legend(reverse = TRUE)) +
  expand_limits(y = c(0, max(nc_camping_cost$cost_avg * 2))) +
  labs(title = "Camping in NC continues to be a great value",
       subtitle = glue("Median cost per person-night",
                    "\nconsidering all NC public campgrounds"),
       x = "",
       y = "")

```

<br>

Camping revenue per park:

```{r}
nc_camping_revenue <- nc_hist %>%
  #filter(str_detect(site_type, "STANDARD") & !str_detect(site_type, "GROUP")) %>%
  filter(!str_detect(site_type, "MANAGEMENT")) %>%
  mutate(year = as.numeric(str_extract(path, "\\d+"))) %>%
  group_by(year, facility_name) %>%
  summarize(n = n(), 
            revenue_in_year = sum(people_nights, na.rm = TRUE)#,
            # cost_median = median(paid_person_night, na.rm = TRUE),
            # cost_avg = mean(paid_person_night, na.rm = TRUE)
            ) %>%
  ungroup()
```

```{r fig.height=8}
facility_name_levels <- nc_camping_revenue %>%
  group_by(facility_name) %>%
  filter(revenue_in_year == max(revenue_in_year),
         n > 1) %>%
  ungroup() %>%
  distinct(facility_name, revenue_in_year) %>%
  arrange(desc(revenue_in_year)) %>%
  pull(facility_name)

nc_camping_revenue %>%
  filter(!str_detect(facility_name, "Picnic"),
         !str_detect(facility_name, "Shelter"),
         n > 1) %>%
  mutate(facility_name = factor(facility_name, levels = facility_name_levels),
  #        # facility_name_label = glue("{facility_name}: {percent(pct_change, accuracy = 1)}"),
  #        # facility_name_label = fct_reorder(facility_name_label, -combined_n_person_nights) #pct_change)
  ) %>%
  ggplot(aes(year, revenue_in_year)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_viridis_d(end = 0.92) +
  guides(color = guide_legend(reverse = TRUE)) +
  expand_limits(y = 0) +
  facet_wrap(~ facility_name, scales = "free_y") +
  labs(title = "Camping revenue by year",
       subtitle = "Subplots ordered by decending maximum yearly camping revenue",
       x = "",
       y = "")

```

<br>

There definitely is missing data. Look at the missing years for Balsam Mountain, Black Mountain, Rocky Bluff and Wolf Ford (among others). Due to changing vendors, all 2018 data for NC is missing. Hanging Dog is accurate--the USDA Forest Service closed it permanently.

My assumption is that camping revenue does not include park entry fees. This is only revenue collected via [recreation.gov](https://recreation.gov).

<br>

Are there trends in the duration of stay for any of the site types?

```{r fig.height=4}
nc_camping_nights <- nc_hist %>%
  filter(year >= 2007) %>% 
  inner_join(., campgrounds_with_2008_data,
             by = "facility_name") %>%
  group_by(year, site_type_simple) %>% 
  summarize(n = n(), 
            n_nights = sum(nights, na.rm = TRUE),
            n_nights_med = median(nights, na.rm = TRUE),
            n_nights_avg = mean(nights, na.rm = TRUE)) %>%
  ungroup()

# campgrounds_with_2008_data <- nc_camping_nights %>%
#   filter(year == 2008) %>%
#   select(facility_name)

# normalization <- nc_camping_nights %>%
#   filter(year == 2008) %>%
#   pull(n_nights)

nc_camping_nights_for_plot <- nc_camping_nights %>%
  
  #complete() %>%
  # group_by(year, site_type_simple) %>% 
  # summarize(n = n(), 
  #           n_nights = sum(n_nights, na.rm = TRUE),
  #           n_nights_med = median(n_nights, na.rm = TRUE),
  #           n_nights_avg = mean(n_nights, na.rm = TRUE)) %>%
  # ungroup() %>%
  
  #mutate(nights_normalized = n_nights / normalization) 
  pivot_longer(cols = starts_with("n"), names_to = "metric", values_to = "value")

nc_camping_nights_for_plot %>%
  filter(!metric %in% c("n", "n_nights"),
         !site_type_simple %in% c("management", "shelter")) %>%
  ggplot(aes(year, value, color = site_type_simple, group = site_type_simple)) +
  geom_line() +
  geom_point(size = 2) +
  #scale_x_continuous(breaks = seq(2005, 2020, by = 5), labels = label_number(accuracy = 1)) +
  #scale_y_continuous(labels = percent_format()) +
  expand_limits(y = 0, x = 2005) +
  theme(plot.title.position = "plot") +
  facet_wrap(~metric) +
  labs(title = "Duration of visits by site type",
       # subtitle = glue("Pecentage of 2008 person-nights; 2008 = {comma(normalization)}",
       #                 "\nn = {nrow(campgrounds_with_2008_data)}; excludes campsites lacking 2008 data"),
       x = "",
       y = "")
```

<br>

Did median group visit really increase so much? Seems like a data error.

Maybe not. Looking at just the last four years of group camping...(not limited to campgrounds with data in 2008)

```{r}
nc_camping_nights <- nc_hist %>%
  filter(year >= 2017,
         site_type_simple == "group") #%>% 
  # inner_join(., campgrounds_with_2008_data,
  #            by = "facility_name") %>%
  # group_by(year, facility_name) %>% 
  # summarize(n = n(), 
  #           n_nights = sum(nights, na.rm = TRUE),
  #           n_nights_med = median(nights, na.rm = TRUE),
  #           n_nights_avg = mean(nights, na.rm = TRUE)) %>%
  # ungroup()
```

```{r}
nc_camping_nights %>%
  filter(nights > 4) %>%
  ggplot(aes(number_of_people, nights, color = year)) +
  geom_jitter(alpha = 0.5) +
  geom_smooth(method = "lm") +
  # scale_x_log10() +
  # scale_y_log10() +
  labs(title = "nights by number of people")
           
```

Where is the order date later than the start date?

```{r}
xx <- nc_hist %>%
  mutate(days_order_to_start = start_date - order_date,
         days_order_to_end = end_date - order_date,
         invalid_start_date = if_else(order_date > start_date + 1, 1, NULL))


#if days_order_to_end > 5 subtract 1 year from order date
```


As you might expect more than 25% of group and cabin reservations are booked six months before the start date.

It looks like in 2007 on could reserve group sites and cabins with 12 months lead time. Then from 2008-2011 the policy changed for cabins: 6 months lead time. Then in 2012 the cabin polic returned to 12 months lead time.

The data for 2019 seems wrong: why are so many order dates after start_date? It' the only year for which that's the case (and the curves in general are depressed for that year).

```{r fig.height=6}
xx %>%
  #mutate(days_order_to_start = if_else(days_order_to_end < -10, days_order_to_start + 365, days_order_to_start)) %>%
  filter(year >= 2007,
         # days_order_to_start > -9,
         # days_order_to_start > -30,
         # days_order_to_start < 366,
         !site_type_simple %in% c("management", "shelter")) %>%
  ggplot(aes(days_order_to_start, color = site_type_simple)) +
  stat_ecdf() +
  coord_cartesian(xlim = c(360, -30)) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format()) +
  geom_vline(xintercept = c(30, 60, 90, 180, 365), color = "firebrick", alpha = 0.5, lty = 2) +
  facet_wrap(~year, ncol = 4) +
  labs(title = "Percent of NC bookings by x days before start_date")
```

<br>

So I adjust for likely data errors, perhaps due to switching vendors in 2019 (and even this adjustment seems like not enough). Best to assume 2018 data is not accurage or complete.

```
# where order_date is more than 10 days after camping end_date, assume the year is off by one
mutate(days_order_to_start = if_else(days_order_to_end < -10, days_order_to_start + 365, days_order_to_start)

```
```{r fig.height=6}
xx %>%
  mutate(days_order_to_start = if_else(days_order_to_end < -10, days_order_to_start + 365, days_order_to_start)) %>%
  filter(year >= 2007,
         # days_order_to_start > -9,
         # days_order_to_start > -30,
         # days_order_to_start < 366,
         !site_type_simple %in% c("management", "shelter")) %>%
  ggplot(aes(days_order_to_start, color = site_type_simple)) +
  stat_ecdf() +
  coord_cartesian(xlim = c(360, -30)) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format()) +
  geom_vline(xintercept = c(30, 60, 90, 180, 365), color = "firebrick", alpha = 0.5, lty = 2) +
  
  facet_wrap(~year, ncol = 4) +
  labs(title = "Percent of NC bookings by x days before start_date")
```

<br>

```{r eval=FASE}
# looking at days_order_to_end
xx %>%
  filter(days_order_to_end > -9,
         days_order_to_end < 366) %>%
  ggplot(aes(days_order_to_end, color = site_type)) +
  stat_ecdf() +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format()) +
  geom_vline(xintercept = c(30, 60, 90, 180, 365), color = "firebrick", alpha = 0.5, lty = 2) +
  labs(title = "Percent of bookings by x days before end_date")

```

# do we have good data for all years (all USA)?

```{r fig.height=6}
#facilities (facilityID and parentorgid)
facilities_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/Facilities_API_v1.json",
                      flatten = FALSE)

facilities <- facilities_raw[[1]] %>%
  clean_names() %>%
  mutate(facility_name = str_to_title(facility_name),
         facility_name = if_else(facility_id == "250838", "Halfway Group Campground", facility_name)
         )

# facility_ids <- facilities %>%
#   distinct(facility_id, .keep_all = TRUE)
```


```{r fig.height=6}
facility_name_years_raw <- read_feather("./data/processed/park-years.feather") %>%
  distinct(facility_id, site_type, start_year, .keep_all = TRUE) %>%
  mutate_if(is.character, ~map_chr(.x, iconv, "UTF-8", "UTF-8", sub='')) %>% # just in case
  left_join(.,
            facilities,
            by = "facility_id") %>%
  mutate(use_type = str_to_lower(use_type),
         facility_name = str_to_lower(facility_name))
```


```{r fig.height=6}
facility_name_years <- left_join(facilities %>% select(facility_id, facility_name, facility_type_description),
                facility_name_years_raw,
                by = "facility_id") %>%
  filter(!is.na(path),
         use_type == "overnight") %>%
  distinct(facility_id, site_type, start_year, .keep_all = TRUE)


# facility_id_hist <- facility_name_years %>%
#   distinct(facility_id, .keep_all = TRUE)
```


The new vendor starting in 2019 allows  more variation in the "park" string. Using the canonical `facility_name` below gives better results. We see the majority of the sites had data in the system in 2008 (not 2007), and across the USA each year between 20 and 50 parks added a new kind of camping site each year. This is consistent with park development.

```{r fig.height=6}
# to avoid duplicates, focus either on start_year or order_year
parks_all_years_site_types <- park_years %>%
  mutate(
    site_type_simple = case_when(
    str_detect(site_type, "TENT")        ~ "tent",
    str_detect(site_type, "CABIN")       ~ "cabin",
    str_detect(site_type, "GROUP")       ~ "group",
    str_detect(site_type, "RV")          ~ "rv",
    str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
    str_detect(site_type, "STANDARD")    ~ "standard",
    str_detect(site_type, "SHELTER")     ~ "shelter",
    str_detect(site_type, "MANAGEMENT")  ~ "management",
    TRUE                                 ~ site_type
  )
  ) %>%
  distinct(facility_id, site_type_simple, start_year, .keep_all = TRUE) %>%
  group_by(facility_id, facility_name, site_type_simple) %>%
  summarize(n_years = n(),
            year_min = min(start_year),
            year_max = max(start_year)
  ) %>%
  ungroup()

parks_all_years_overnight <- park_years %>%
  filter(use_type == "overnight") %>%
  mutate(
    site_type_simple = case_when(
      str_detect(site_type, "TENT")        ~ "tent",
      str_detect(site_type, "CABIN|YURT")  ~ "cabin",
      str_detect(site_type, "GROUP")       ~ "group",
      str_detect(site_type, "RV")          ~ "rv",
      str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
      str_detect(site_type, "STANDARD")    ~ "standard",
      str_detect(site_type, "SHELTER")     ~ "shelter",
      str_detect(site_type, "MANAGEMENT")  ~ "management",
      str_detect(site_type, "HIKE TO|WALK IN|BOAT TO")  ~ "tent", # assume these are tent
      TRUE                                 ~ "other"
    )
  ) %>%
  distinct(facility_id, site_type_simple, start_year, .keep_all = TRUE) %>%
  group_by(facility_id, facility_name, site_type_simple) %>%
  summarize(n_years = n(),
            year_min = min(start_year, na.rm = TRUE),
            year_max = max(start_year, na.rm = TRUE)
  ) %>%
  ungroup()

# parks_all_years_overnight %>%
#   count(site_type_simple, sort = TRUE)
# TODO: why is this not summarizing by site_type_simple?

parks_all_years_overnight_for_plot <- parks_all_years_overnight %>%
  filter(site_type_simple != "management") %>%
  mutate(site_type_simple = fct_reorder(site_type_simple, -n_years))
```

<br>

```{r fig.height=4}
parks_all_years_overnight_for_plot %>%
  count(year_min, site_type_simple) %>%
  ggplot(aes(x = year_min, y = n, fill = site_type_simple)) +
  geom_area(stat = "identity") +
  scale_y_continuous(label = comma_format()) +
  scale_fill_viridis_d(end = 0.92, direction = -1) +
  labs(title = "Number of facilities newly offering site_type_simple with 'Overnight' use_type by year",
       subtitle = glue("As represented in published recreation.gov data set",
                       " {min(parks_all_years_overnight_for_plot$year_min)}", 
                       " to {max(parks_all_years_overnight_for_plot$year_min)}",
                       "\nBy 2008 it seems existing sites were reservable; later years are likely organic growth"),
       x = "Year newly offering reservations for each kind of site offer overnight reservations at recreation.gov",
       y = "Number of facilities in the USA"
  )
```

<br>

```{r fig.height=6}
year_lower_bound <- 2009

parks_all_years_overnight_for_plot %>%
  filter(year_min > year_lower_bound) %>%
  ggplot(aes(year_min, fill = site_type_simple)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020)) +
  facet_wrap(~site_type_simple) +
  scale_fill_viridis_d(end = 0.92, direction = -1) +
  labs(title = "Number of facilities newly offering site_type_simple with 'Overnight' use_type by year (2009+)",
       subtitle = glue("As represented in published recreation.gov data set",
                       " {year_lower_bound}", 
                       " to {max(parks_all_years_overnight_for_plot$year_min)}"),
       x = "Year newly offering reservations for each kind of site offer overnight reservations at recreation.gov",
       y = "Number of facilities in the USA"
  )
```

<br>

CAGRs seem reasonable too: (TODO: check this!!)

```{r fig.height=4}
parks_all_years_overnight <- park_years %>%
  filter(use_type == "overnight") %>%
  mutate(
    site_type_simple = case_when(
      str_detect(site_type, "TENT")        ~ "tent",
      str_detect(site_type, "CABIN|YURT")       ~ "cabin",
      str_detect(site_type, "GROUP")       ~ "group",
      str_detect(site_type, "RV")          ~ "rv",
      str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
      str_detect(site_type, "STANDARD")    ~ "standard",
      str_detect(site_type, "SHELTER")     ~ "shelter",
      str_detect(site_type, "MANAGEMENT")  ~ "management",
      str_detect(site_type, "HIKE TO|WALK IN|BOAT TO")  ~ "tent", # assume these are tent
      TRUE                                 ~ "other"
    )
  ) %>%
  distinct(facility_id, site_type_simple, start_year, .keep_all = TRUE) %>%
  group_by(start_year, site_type_simple) %>%
  summarize(n_parks = n()) %>%
  ungroup()

#from https://datascience.stackexchange.com/questions/26347/compound-annual-growth-rate-function-in-r-statistics

CAGR_formula <- function(FV, PV, yrs) {
  (FV/PV)^(1/yrs)-1
}

parks_all_years_overnight_for_plot <- parks_all_years_overnight %>%
  filter(site_type_simple != "management",
         start_year >= 2008) %>%
  group_by(site_type_simple) %>%
  #mutate(agr_by_site_type = round(100 * max(n_parks / min(n_parks)) / (max(start_year) - min(start_year)) + 1), 0) %>%
  mutate(agr_by_site_type = sum(n_parks[start_year == 2020]) / sum(n_parks[start_year == 2008]) / (2020 - 2008),
         cagr_by_site_type = CAGR_formula(sum(n_parks[start_year == 2020]), 
                                          sum(n_parks[start_year == 2008]), 
                                          2020 - 2008)
         ) %>%
  ungroup() %>%
  mutate(site_type_simple_label = glue("{site_type_simple} ({percent(cagr_by_site_type, accuracy = 0.1)})"),
    site_type_simple_label = fct_reorder(site_type_simple_label, -n_parks)) 


agr <- with(parks_all_years_overnight_for_plot,
            round(sum(n_parks[start_year == 2020]) / sum(n_parks[start_year == 2008])) / (2020 - 2008 + 1), 0)


cagr <- with(parks_all_years_overnight_for_plot,
             CAGR_formula(sum(n_parks[start_year == 2020]), sum(n_parks[start_year == 2008]), yrs = 2020 - 2008 + 1)
             )
  
  
parks_all_years_overnight_for_plot %>%
  ggplot(aes(start_year, n_parks, color = site_type_simple_label)) +
  geom_line() +
  geom_point(size = 2) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020)) +
  #facet_wrap(~site_type_simple) +
  scale_color_viridis_d(end = 0.92, direction = -1) +
  labs(title = "Number of facilities offering site_type_simple with 'Overnight' use_type by year",
       subtitle = glue(#"combined annual growth rate (AGR): {percent(last(agr), accuracy = 0.1)}"),
                       "Compound annual growth rate (CAGR) for all offering site types: {percent(cagr, accuracy = 0.1)}",
                       " from {min(parks_all_years_overnight_for_plot$start_year)} to {max(parks_all_years_overnight_for_plot$start_year)}",
                       "\nAs represented in published recreation.gov data set"),
       x = "Reservation start year",
       color = "Site type (CAGR)")
```

<br>

```{r fig.height=5}
library(ggridges)

# parks_all_years_overnight %>%
#   mutate(start_year = as.character(start_year),
#          park_site_type_unit = 1) %>%
#   group_by(start_year, site_type_simple) %>%
#   mutate(n_parks = n(),
#          n_park_with_site_type = sum(park_site_type_unit)) %>%
#   ungroup() %>%
  #mutate(site_type_simple = fct_reorder(site_type_simple, -n_parks)) %>%


park_years %>%
  # filter(use_type == "overnight",
  #        !str_detect(site_type, "MANAGEMENT")) %>%
  # mutate(
  #   site_type_simple = case_when(
  #     str_detect(site_type, "TENT")        ~ "tent",
  #     str_detect(site_type, "CABIN|YURT")       ~ "cabin",
  #     str_detect(site_type, "GROUP")       ~ "group",
  #     str_detect(site_type, "RV")          ~ "rv",
  #     str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
  #     str_detect(site_type, "STANDARD")    ~ "standard",
  #     str_detect(site_type, "SHELTER")     ~ "shelter",
  #     str_detect(site_type, "MANAGEMENT")  ~ "management",
  #     str_detect(site_type, "HIKE TO|WALK IN|BOAT TO")  ~ "tent", # assume these are tent
  #     TRUE                                 ~ "other"
  #   )
  # ) %>%
  # distinct(facility_id, site_type_simple, start_year, .keep_all = TRUE) %>%
  ggplot() + #, fill = site_type_simple)) +
  geom_density_ridges(aes(y = site_type_simple, x = start_year, fill = site_type_simple),
                      alpha = 0.6) +
  #facet_wrap(~site_type_simple) +
  labs(title = "Number of facilities offering site_type_simple with 'Overnight' use_type by year",
       subtitle = "As represented in published recreation.gov data set")
```

<br>

```{r}
park_years %>%
  filter(use_type == "overnight",
         !str_detect(site_type, "MANAGEMENT")) %>%
  mutate(
    site_type_simple = case_when(
      str_detect(site_type, "TENT")        ~ "tent",
      str_detect(site_type, "CABIN|YURT")       ~ "cabin",
      str_detect(site_type, "GROUP")       ~ "group",
      str_detect(site_type, "RV")          ~ "rv",
      str_detect(site_type, "EQUESTRIAN")  ~ "equestrian",
      str_detect(site_type, "STANDARD")    ~ "standard",
      str_detect(site_type, "SHELTER")     ~ "shelter",
      str_detect(site_type, "MANAGEMENT")  ~ "management",
      str_detect(site_type, "HIKE TO|WALK IN|BOAT TO")  ~ "tent", # assume these are tent
      TRUE                                 ~ "other"
    )
  ) %>%
  distinct(facility_id, site_type_simple, start_year, .keep_all = TRUE) %>%
  group_by(start_year, site_type_simple) %>%
  summarize(n_parks = n()) %>%
  ungroup() %>%
  mutate(n_parks_scaled = n_parks / max(n_parks),
         site_type_simple = fct_reorder(site_type_simple, n_parks, max)) %>%
  ggplot() + #, fill = site_type_simple)) +
  geom_ridgeline(aes(y = site_type_simple, x = start_year, height = n_parks_scaled, fill = site_type_simple),
                      alpha = 0.6) +
  scale_fill_viridis_d(end = 0.92, direction = 1) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(legend.position = "none") +
  #facet_wrap(~site_type_simple) +
  labs(title = "Number of facilities offering site_type_simple \nwith 'Overnight' use_type by year",
       subtitle = "As represented in published recreation.gov data set",
       x = NULL,
       y = NULL)
```

