---
title: "load-data (experimental)"
author: "Daniel Moul"
date: "`r Sys.Date()`"
output: html_document
---

<br><br>

***This is an experimental script used prior to the development of nc-camping-writeup.Rmd. It is not needed and may not work correctly. If you do use it, run individual chunks***

<br>

```{r}
# Stop someone from running this naively
knitr::knit_exit()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse, warn.conflicts = FALSE)
library(arrow, warn.conflicts = FALSE)
library(jsonlite)
library(janitor)
library(gt)
library(scales)
library(glue)

library(hrbrthemes)
# remotes::install_github("hrbrmstr/hrbrthemes")
# run 2021-08-01 to address segfault errors like this:
# sh: line 1: 11020 Segmentation fault: 11  '/Library/Frameworks/R.framework/Versions/4.1/Resources/library/Rttf2pt1/exec//ttf2pt1' -a -GfAe '/Library/Frameworks/R.framework/Versions/4.1/Resources/library/hrbrthemes/fonts/roboto-condensed/RobotoCondensed-Bold.ttf' '/var/folders/0j/hjw8sflx19l6rw4s1l68qq640000gn/T//RtmpGN3PSJ/fonts/RobotoCondensed-Bold' 2>&1
theme_set(theme_ipsum_ps() + 
            theme(panel.grid.minor = element_blank()))
import_plex_sans()

# library(readr)
# options(readr.read_lazy = TRUE)

options(dplyr.summarise.inform = FALSE)

big_campground_size <- 50 # fifty or more campsites

```

```{r}
#rec_areas <- read_json_arrow("./data/raw/RIDBFullExport_V1_JSON/RecAreas_API_v1.json")
rec_areas_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/RecAreas_API_v1.json",
                      flatten = FALSE)
rec_areas <- rec_areas_raw[[1]] %>%
  clean_names()

# campsites (campsidetid and facilityid)
campsites_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/Campsites_API_v1.json",
                      flatten = FALSE)
campsites <- campsites_raw[[1]] %>%
  clean_names()

#attributes per campsite (entityid)
# campsite_attributes_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/CampsiteAttributes_API_v1.json",
#                       flatten = FALSE)
# campsite_attributes <- campsite_attributes_raw[[1]] %>%
#   clean_names()

# activityid codes
activities_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/Activities_API_v1.json",
                      flatten = FALSE)
activities <- activities_raw[[1]] %>%
  clean_names()

#facilities (facilityID and parentorgid)
facilities_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/Facilities_API_v1.json",
                      flatten = FALSE)
facilities <- facilities_raw[[1]] %>%
  clean_names()

#facilities (facilityID and parentorgid)
facility_addresses_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/FacilityAddresses_API_v1.json",
                      flatten = FALSE)
facility_addresses <- facility_addresses_raw[[1]] %>%
  clean_names()
# postal_code and address_state_code are reversed, and some states like UTAH are not always abbreviated

nc_facility_addresses <- facility_addresses %>%
  filter(postal_code == "NC")


orgs_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/Organizations_API_v1.json",
                      flatten = FALSE)
orgs_for_joining <- orgs_raw[[1]] %>%
  clean_names() %>%
  select(org_id, org_name, org_abbrev_name, org_url_address)

# this isn't helpful to me
# orgs_entities_raw <- fromJSON("./data/raw/RIDBFullExport_V1_JSON/OrgEntities_API_v1.json",
#                       flatten = FALSE)
# orgs_entities <- orgs_entities_raw[[1]] %>%
#   clean_names()

```

<br>

There are `r nrow(nc_facility_addresses)` facility addresses in NC. Which have camping? Presumably only those with physical addresses:

<br>

```{r}

nc_facility_addresses %>% count(facility_address_type) %>%
  gt()
```

<br><br>

Are there any duplicated facility IDs? Yes: a small number:

<br>

```{r}
nc_facility_addresses %>% count(facility_id) %>%
  filter(n > 1) %>%
  mutate(facility_id = as.numeric(facility_id)) %>%
  arrange(-n, facility_id) %>%
  gt()

```

<br><br>

It seems where there are duplicates, each mailing address has at least one physical address:

<br>

```{r}
nc_facility_addresses %>% 
  group_by(facility_id) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  mutate(facility_id = as.numeric(facility_id)) %>%
  arrange(-n, facility_id) %>%
  gt()
```

<br><br>

So let's assume that all campsites will have physical or default addresses.

```{r}
nc_facility_ids_for_join <- nc_facility_addresses %>%
  filter(facility_address_type != "Mailing") %>%
  distinct(facility_id) # there is one with two physical addresses
  
nc_facilities <- inner_join(nc_facility_ids_for_join,
                            facilities,
                            by = "facility_id") %>%
  mutate(facility_name = str_to_title(facility_name))

nc_campsites <- inner_join(nc_facility_ids_for_join,
                           campsites,
                           by = "facility_id") %>%
  mutate(campsite_type = str_to_lower(campsite_type))

nc_campsite_ids_for_join <- nc_campsites %>%
  distinct(campsite_id)

nc_facilities_with_campsites <- inner_join(nc_facilities,
                                           nc_campsites,
                                           by = "facility_id")
# One row per campsite
# I note there are some USDA sites without lon/lat


nc_campground_n_sites <- nc_facilities_with_campsites %>%
  group_by(facility_id) %>%
  summarize(n = n()) %>%
  ungroup() 

```

There are `r nrow(distinct(nc_facilities_with_campsites, facility_id))` campgrounds with `r comma(nrow(distinct(nc_facilities_with_campsites, facility_id, campsite_id)))` campsites in total.  The distribution of number of campsites per campground:

```{r}
nc_campground_n_sites %>% 
  ggplot(aes(n)) +
  stat_ecdf(size = 2, color = "firebrick") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = glue("Percent of all {nrow(distinct(nc_facilities_with_campsites, facility_id))}", 
                    " NC campgrounds with n campsites"),
       subtitle = glue("{nc_campground_n_sites %>% filter(n ==1) %>% nrow()}",
                     " campgrounds have only one site (typically a group site)"),
       y = "")

```

<br>

Campgrounds with at least 100 campsites:

```{r}
nc_campground_n_big_sites_for_table <- nc_campground_n_sites %>%
  filter(n >= big_campground_size) %>%
  inner_join(.,
             nc_facilities,
             by = "facility_id") %>%
  # TODO: could modify the commented code to add subcatery: type of campsite
  # inner_join(.,
  #            nc_campsites,
  #            by = "facility_id") %>%
  select(facility_id, facility_name, n_campsites = n) %>%
  arrange(-n_campsites) 

nc_campground_n_big_sites_for_table %>%
  gt() %>%
  tab_header(md(glue("**{nrow(nc_campground_n_sites_for_table)} NC campgrounds with at least {big_campground_size} campsites**")))
```

<br><br>

```{r}
nc_campground_n_all_sites_for_table <- nc_campground_n_sites %>%
  inner_join(.,
             nc_facilities,
             by = "facility_id") %>%
  # TODO: could modify the commented code to add subcatery: type of campsite
  # inner_join(.,
  #            nc_campsites,
  #            by = "facility_id") %>%
  select(facility_id, facility_name, n_campsites = n) %>%
  arrange(-n_campsites) 

nc_campground_n_all_sites_for_table %>%
  inner_join(.,
             nc_facilities,
             by = "facility_id") %>%
  inner_join(.,
             nc_campsites %>%
               group_by(facility_id, campsite_type) %>%
               summarize(campsite_type_n = n()) %>%
               ungroup(),
             by = "facility_id") %>%
  select(facility_id, facility_name, n_campsites = n, campsite_type, campsite_type_n) %>%
  arrange(-n_campsites, -campsite_type_n) %>%
  mutate(facility_label = glue("{facility_id} - {facility_name}: {n_campsites} sites")) %>%
  select(-c(facility_id, facility_name, n_campsites)) %>%
  group_by(facility_label) %>%
  gt() %>%
  tab_header(md(glue("**NC campgrounds with at least {big_campground_size} campsites**<br>With count by campsite type")))
  

```

<br><br>

```{r}
nc_campground_n_sites %>%
  filter(n >= big_campground_size) %>%
  inner_join(.,
             nc_facilities,
             by = "facility_id") %>%
  inner_join(.,
             nc_campsites %>%
               group_by(facility_id, campsite_type) %>%
               summarize(campsite_type_n = n()) %>%
               ungroup(),
             by = "facility_id") %>%
  select(facility_id, facility_name, n_campsites = n, campsite_type, campsite_type_n) %>%
  arrange(-n_campsites, -campsite_type_n) %>%
  mutate(facility_label = glue("{facility_id} - {facility_name}: {n_campsites} sites"),
         facility_label = fct_reorder(facility_label, n_campsites)) %>%
  select(-c(facility_id, facility_name, n_campsites)) %>%
  mutate(campsite_type = case_when(
    str_detect(campsite_type, "group")     ~ "group or shelter",
    str_detect(campsite_type, "shelter")   ~ "group or shelter",
    TRUE                                   ~ campsite_type
  ),
  campsite_type = fct_reorder(campsite_type, campsite_type_n, sum)
  ) %>%
  group_by(facility_label) %>%
  ggplot(aes(campsite_type_n, facility_label, fill = campsite_type)) +
  geom_col() +
  scale_fill_viridis_d(direction = -1) +
  theme(plot.title.position = "plot") +
  labs(title = glue("NC campgrounds with at least {big_campground_size} campsites"),
       subtitle = "By campsite type",
       y = "")

```

