---
title: "load-historical-data"
author: "Daniel Moul"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse, warn.conflicts = FALSE)
library(arrow, warn.conflicts = FALSE)
library(jsonlite)
library(janitor)
library(gt)
library(scales)
library(glue)
library(lubridate)
library(purrr)

library(hrbrthemes)
theme_set(theme_ipsum_ps() + 
            theme(panel.grid.minor = element_blank()))
import_plex_sans()

# library(readr)
# options(readr.read_lazy = TRUE)

options(dplyr.summarise.inform = FALSE)

big_campground_size <- 50 # fifty or more campsites

```

<br>

<!-- Let's start by looking at 2020 only. -->

```{r eval=FALSE}

res2020 <- read_csv("./data/raw/FY20 Historical Reservations Full.csv")

nc_res_2020 <- filter(res2020, facilitystate == "North Carolina") %>%
  select(-c(historicalreservationid, ordernumber, codehierarchy, legacyfacilityid, tax:totalpaid, orderdate,
            regioncode, regiondescription),
            -starts_with(c("customer", "equipment"))) %>%
  rename(facility_id = facilityid,
         org_id = orgid,
         campsite_id = productid) %>%
  filter(inventorytype == "CAMPING",
         usetype == "Overnight"#,
         #str_detect(nights, "day") 
         ) %>%
  separate(nights, into = c("nights", "units"), fill = "right", extra = "drop") %>%
  filter(str_detect(units, "day")) %>% # drop ~ 100 rows with negative days or 00 days
  mutate(nights = as.numeric(nights),
         park = str_to_title(park),
         sitetype = str_to_lower(sitetype)) %>%
  select(-units)
  

# product_id here == campsite_id in campsites df

#xx <- names(nc_res_2020)


```

<!-- How many rows have the same productid aka campsite_id values? Theoretical max is 365 if each reservation was for one night. Looking at the ten most frequent productid values, clearly at least the top three rows are using the same productid for multiple sites. -->

```{r eval=FALSE}
count(nc_res_2020, park, sitetype, campsite_id, sort = TRUE) %>%
  slice_max(order_by = n, n = 10) %>%
  gt()
```

<br>

do we have the same fields in all historical csv files?

```{r}
path_csv <- "./data/raw/reservations/"

all_files_csv_hist <- fs::dir_ls(path = path_csv, # root_dir,
                        recurse = FALSE,
                        type = "file",
                        regexp = "[.]csv$"
)

get_csv_headers <- function(fname, n_rows = 1) {
  read_csv(file = fname, 
           n_max = n_rows,
           id = "path",
           col_names = FALSE)
}

csv_historical_file_columns <- purrr::map_dfr(all_files_csv_hist, get_csv_headers)

```

HMMM data changes over time (and this doesn't include data type changes)

```{r}
csv_historical_file_columns %>%
  gt()

```


Actions (done)

* 2019 and 2020 differ the most (and slightly from each other)
    - use col names from earlier years; they will parse better with clean_names() and better align with facility & campground data that came in JSON
    - region_code and region_description may not be the same info
    - though 2017, entity_type == "Site" is camping; equivalent to 2019+2010 inventory_type = "CAMPING"
    - though 2017, entity_id seems to be extra--there is no equivalent in 2019+2010 product_id
    - through 2017 facility_state is like "NC" but in 2019+2020 it's like "North Carolina"
* 2019: calculate `nights`
* all: drop cols to the right of NumberOfPeople
* 

```{r define-nc-by-year-feather}

# the following works, but it was only used to figure out the functions (not needed now)
# nc_2006 <- read_csv("./data/raw/reservations/2006.csv",
#                     col_types = cols(.default = "c"),
#                     id = "path") %>%
#   clean_names() %>%
#   filter(facility_state == "NC") %>%
#   mutate(across(contains("_date"), ymd)) %>%
#   mutate(across(ends_with(c("tude", "_people")), as.numeric)) %>%
#   select(1:number_of_people) %>%
#   mutate(nights = as.numeric(difftime(end_date, start_date, units = "days"))) %>%
#   filter(nights > 0) # drop day use shelters
  
get_nc_historical <- function(fname) {
  
  # test
  # fname <- "./data/raw/reservations/2010.csv"
  
  tbl_tmp <- read_csv(file = fname, 
                      col_types = cols(.default = "c"),
                      id = "path") %>%
    clean_names()
    
  tbl_tmp2 <- tbl_tmp %>%
    filter(facility_state == "NC") %>%
    mutate(across(ends_with("date"), ~ as.Date(.x, origin = "1970-01-01"))) %>%
    mutate(across(ends_with(c("tude", "_people")), as.numeric)) %>%
    select(1:number_of_people) %>%
    # mutate(nights = as.numeric(end_date - start_date)) %>%
    mutate(nights = as.numeric(difftime(end_date, start_date, units = "days"))) %>%
    filter(nights > 0) # drop day use shelters
  
  tbl_tmp2
}


get_nc_historical_2019plus <- function(fname) {
  
  # test
  # fname <- "./data/raw/reservations/2020.csv"
  
  tbl_tmp <- read_csv(file = fname, 
           col_types = cols(.default = "c"),
           id = "path") %>%
    clean_names() 
  
 tbl_tmp2 <- tbl_tmp %>%
    filter(facilitystate == "North Carolina", # drop day use shelters
           inventorytype != "VEHICLE_PERMIT") %>% # not off-road permits
    mutate(across(ends_with("date"), ~ as.Date(.x, origin = "1970-01-01"))) %>%
    mutate(across(ends_with(c("tude", "people")), as.numeric)) %>%
    select(1:numberofpeople, - starts_with(c("entityid", "equipment", "discount"))) %>% #some don't exist pre-2019
    mutate(nights = as.numeric(difftime(enddate, startdate, units = "days"))) %>%
    filter(nights > 0) 
  
  # entity_type replaces inventorytype; removed "entity_id", 
  names(tbl_tmp2) <- c("path", "historical_reservation_id", "order_number", "agency", "org_id",
                      "code_hierarchy", "region_code", "region_description", "parent_location_id", "parent_location",
                      "legacy_facility_id", "park", "site_type", "use_type", "product_id",
                      "entity_type", "facility_id", "facility_zip", "facility_state",
                      "facility_longitude", "facility_latitude", "customer_zip", "customer_state", "customer_country",
                      "tax", "use_fee", "tran_fee", "attr_fee", "total_before_tax",
                      "total_paid", "start_date", "end_date", "order_date", "number_of_people", "nights")
  
  tbl_tmp2
}


# nc_historical_part1 <- purrr::map_dfr(all_files_csv_hist[str_detect(all_files_csv_hist, "2006|2007|2008|2009|2010|2011|2012|2013|2014|2015|2016|2017|2018")],
                                # get_nc_historical)
# nc_historical_part2 <- purrr::map_dfr(all_files[str_detect(all_files, "2019|2020")],
#                                 get_nc_historical_2019plus)
```

```{r define-nc_hist}

path_feather <- "./data/processed/nc_by_year_semiprocessed"

all_files <- fs::dir_ls(path = path_feather, #root_dir,
                        recurse = FALSE,
                        type = "file",
                        regexp = "[.]feather$"
)


nc_hist_tmp <- map_dfr(all_files, read_feather)

nc_hist <- nc_hist_tmp %>%
  select(-c(starts_with(c("historical", "legacy")), region_code, code_hierarchy, starts_with("parent"))) %>%
  mutate(entity_type = if_else(entity_type == "Site", "CAMPING", entity_type),
         across(tax:total_paid, as.numeric),
         people_nights  = number_of_people * nights,
         paid_person_night = total_paid / people_nights
  )

write_feather(nc_hist, paste0("./data/processed/nc_historical_camping.feather"))

```

How many reservations per year?

```{r}
nc_hist %>%
  group_by(year(start_date)) %>%
  summarize(n = n()) %>%
  adorn_totals(where = "row") %>%
  gt()
```

```{r}
nc_hist %>%
  group_by(path) %>%
  summarize(n = n()) %>%
  adorn_totals(where = "row") %>%
  gt()
```

<br>

Check data completeness--all states

```{r define-nc_hist}

path_csv <- "./data/raw/reservations/"

all_files_csv_hist <- fs::dir_ls(path = path_csv, # root_dir,
                        recurse = FALSE,
                        type = "file",
                        regexp = "[.]csv$"
)

get_park_years <- function(fname) {
  
  # test
  # fname <- "./data/raw/reservations/2020.csv"
  
  tbl_tmp <- read_csv(file = fname, 
                      col_types = cols(.default = "c"),
                      id = "path") %>%
    clean_names()
  
  tbl_tmp2 <- tbl_tmp
  
  if(any(str_detect(tbl_tmp2$path, "2019|2020"))) {
    tbl_tmp2 <- tbl_tmp2 %>%
      select(-starts_with(c("entityid", "equipment", "discount")))
    
    if(any(str_detect(tbl_tmp2$path, "2020"))) {
      tbl_tmp2 <- tbl_tmp2 %>% 
        select(-nights)
    }

    names(tbl_tmp2) <- c("path", "historical_reservation_id", "order_number", "agency", "org_id",
                         "code_hierarchy", "region_code", "region_description", "parent_location_id", "parent_location",
                         "legacy_facility_id", "park", "site_type", "use_type", "product_id",
                         "entity_type", "facility_id", "facility_zip", "facility_state",
                         "facility_longitude", "facility_latitude", "customer_zip", "customer_state", "customer_country",
                         "tax", "use_fee", "tran_fee", "attr_fee", "total_before_tax",
                         "total_paid", "start_date", "end_date", "order_date", "number_of_people") #, "nights") #TODO: are last two switched?
    
  }
  
  tbl_tmp2 <- tbl_tmp2 %>%
    select(path, agency, org_id, park:facility_latitude, ends_with("date")) %>%
    mutate(across(ends_with("date"), ~ as.Date(.x, origin = "1970-01-01"))) %>%
    mutate(across(ends_with(c("tude", "_people")), as.numeric)) %>%
    #select(1:number_of_people) %>%
    # mutate(nights = as.numeric(end_date - start_date)) %>%
    mutate(start_year = year(start_date),
           order_year = year(order_date)) %>%
    distinct(facility_id, site_type, start_year, order_year, .keep_all = TRUE)
  
  tbl_tmp2
}

# all_files_selected <- all_files_csv_hist[str_detect(all_files_csv_hist, as.character(2019))]

# TODO: Run this again when I go to dinner
park_years_raw <- map_dfr(all_files_csv_hist, get_park_years)

park_years <- park_years_raw %>%
  distinct(facility_id, site_type_simple, start_year, order_year, .keep_all = TRUE)

write_feather(park_years, paste0("./data/processed/park-years.feather"))

```