# Camping cost per person-night

```{r}
#| label: "setup"
#| include: FALSE

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(here)
source(here("scripts", "my-setup.R"))

```

```{r}
#| label: "read-target-objects"

# facilities <- tar_read(facilities_all)
# federal_org_info <- tar_read(federal_orgs)

nc_history <- tar_read(nc_camping_history)

year_min <- year_lower_bound
year_max <- max(nc_history$year)

# # one row for each overnight stay (includes day of week (dow))
# nc_history_detail <- tar_read(nc_camping_history_detail) %>%
#   mutate(year = year(start_date))

# nc_parks_site_types <- nc_history %>%
#   mutate(
#       name_first_word = str_extract(facility_name, "\\w+")
#   ) %>% 
#   # slice_sample(n = 1e5) %>% # for testing
#   # This works because each first word is unique--it probably wouldn't work if looking at whole USA
#   group_by(name_first_word) %>%
#   mutate(facility_id_group = cur_group_id()
#   ) %>%
#   ungroup() %>%
#   group_by(facility_id_group, site_type) %>%
#   mutate(n_sites = n_distinct(product_id)) %>%
#   ungroup() %>%
#   distinct(facility_id_group, facility_id, site_type, .keep_all = TRUE) %>% 
#   select(name_first_word, facility_id, facility_name, facility_type_description, site_type, facility_id_group, n_sites) %>%
#   group_by(facility_id_group) %>%
#   mutate(facility_name = as.list(facility_name),
#          facility_name = glue_collapse(unique(facility_name), sep = " + "),
#          facility_id_list = as.list(facility_id),
#          facility_id_list = glue_collapse(unique(facility_id_list), sep = ", ")
#   ) %>%
#   ungroup() %>%
#   arrange(facility_name, site_type)
# # one row for each facility_name (combined) and site_type
# # for cases where only the facility_name ang or facility_group_id matters, use distinct()


# nc_parks_site_types_product_id <- nc_history %>%
#   mutate(
#       name_first_word = str_extract(facility_name, "\\w+")
#   ) %>% 
#   # slice_sample(n = 1e5) %>% # for testing
#   group_by(name_first_word) %>%
#   mutate(facility_id_group = cur_group_id()
#   ) %>%
#   ungroup() %>%
#   group_by(facility_id_group, site_type) %>%
#   mutate(product_id_list = list(unique(product_id)),
#          n_sites = length(product_id_list)) %>%
#   ungroup() %>%
#   distinct(facility_id_group, facility_id, site_type, .keep_all = TRUE) %>% 
#   select(name_first_word, facility_id, facility_name, facility_type_description, site_type, facility_id_group, n_sites) %>%
#   group_by(facility_id_group) %>%
#   mutate(facility_name = as.list(facility_name),
#          facility_name = glue_collapse(unique(facility_name), sep = " + "),
#          facility_id_list = as.list(facility_id),
#          facility_id_list = glue_collapse(unique(facility_id_list), sep = ", ")
#   ) %>%
#   ungroup() %>%
#   arrange(facility_name, site_type)
# # one row for each facility_name (combined) and site_type
# # for cases where only the facility_name or facility_group_id matters, use distinct()

# nc_campsites <- tar_read(nc_campsites_from_history) %>%
#   rename(facility_name_original = facility_name) %>%
#   left_join(.,
#             nc_parks_site_types %>%
#               select(facility_id_group, facility_name, facility_id = facility_id_list) %>%
#               separate_rows(facility_id),
#             by = "facility_id"
#             ) %>%
#   distinct(facility_id_group, facility_name, campsite_id)
  
# park_years_nc <- tar_read(park_years_nc)

# # feather file seems to be flakey, so using targets() cached value instead
# # park_years <- readRDS("./data/processed/park-years.rds") %>% # could use .rds if we wanted to
# park_years_usa <- tar_read(park_years_usa) %>%  
#   filter(facility_name != "Washington Monument",
#          !is.na(start_year)) %>%
#   # Fix a data error
#   mutate(
#     facility_latitude = if_else(facility_name == "Cape Point Campground", 35.235278, facility_latitude),
#     facility_longitude = if_else(facility_name == "Cape Point Campground", -75.535278, facility_longitude)
# )

# nc_yday <- tar_read(nc_camping_history_yday) %>%
#   count(year, facility_name, site_type, season, day_of_reservation, yday_of_reservation,
#         name = "n_campsites") %>% # number of campsites used that night (not person_nights)
#   arrange(facility_name, year, yday_of_reservation)

# nc_week <- tar_read(nc_camping_history_yday) %>%
#   unnest(week_of_reservation, week_of_reservation2, yday_of_reservation) %>%
#   count(year, facility_name, site_type, season, week_of_reservation, #n_campgrounds, 
#         name = "n_campsites") %>%
#   arrange(facility_name, year, week_of_reservation, site_type)

# nc_week_n_camps_occupied <- tar_read(nc_camping_history_yday) %>%
#   filter(year >= year_lower_bound,
#          site_type != "management") %>%
#   distinct(year, week_of_reservation, facility_name, site_type) %>%
#   mutate(week_of_reservation = map(week_of_reservation, unique)) %>%
#   unnest(week_of_reservation) %>%
#   distinct(year, week_of_reservation, facility_name, site_type) %>%
#   group_by(year, week_of_reservation, site_type) %>%
#   summarize(n_campgrounds = n_distinct(facility_name)) %>%
#   ungroup()

# nc_week2 <- nc_history_detail %>%
#   mutate(week_of_reservation = epiweek(overnight_date)) %>%
#   group_by(facility_name, site_type, loc_mountains, year, week_of_reservation) %>%
#   summarize(n_campsites = n_distinct(product_id)) %>%
#   ungroup()

# nc_week_n_camps_occupied2 <- nc_week2

```


## What does it cost per person, and how has that changed?

Median cabin cost per person doubled in 2012, and equestrian per-person cost grew 2013- 2017. Otherwise median prices were generally stable. Note that these costs are not inflation-adjusted.

Per-person costs at group and equestrian sites in 2020 seem like an aberration: either the more expensive sites were not used much, group sizes were a lot larger (which seems unlikely during the pandemic), there was some other change in usage pattern, or there could be a problem with the data.

```{r}
#| label: "nc-great-value"
#| fig.width: 9
#| fig.height: 6
#| column: page-right

nc_camping_cost <- nc_history %>%
  filter(year >= year_lower_bound,
         !str_detect(site_type, "management")) %>%
  group_by(year, site_type) %>%
  summarize(n = n(), 
            n_person_nights = sum(person_nights, na.rm = TRUE),
            cost_median = median(paid_person_night, na.rm = TRUE),
            cost_avg = mean(paid_person_night, na.rm = TRUE)) %>%
  ungroup()

nc_camping_cost %>%
  filter(site_type != "shelter") %>%
  mutate(site_type = fct_reorder(site_type, cost_median, last)) %>%
 ggplot(aes(year, cost_median, color = site_type, size = n_person_nights)) +
  geom_line(alpha = 0.7, show.legend = FALSE) +
  geom_point(size = rel(6)) + #
  scale_x_continuous(breaks = seq(2005, 2020, by = 5)) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_viridis_d(end = 0.92) +
  scale_size_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size = 3))) +
  expand_limits(y = 0) +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Camping in NC continues to be a great value",
       subtitle = glue("Median cost per person-night",
                    "\nat NC public campgrounds"),
       x = "",
       y = "",
       color = "Site type",
       size = "Number of\nperson nights",
       caption = my_caption)

```

<br>
